---
title: "IBD Literature Report"
subtitle: "Week of `r format(Sys.Date(), '%B %d, %Y')`"
date: last-modified
execute:
  echo: false
  warning: false
---

```{r setup}
#| include: false
library(tidyverse)
library(glue)
library(knitr)

# Source core functions
source("R/core_functions.R")

# Get papers (uses cache if available)
papers <- get_papers(days_back = 30, use_cache = TRUE)

# Generate RSS feed
generate_rss_feed(papers)
```

::: {.callout-note icon=false}
## IBD LitMonitor

All IBD papers this week, **organized by subfield**. Every paper, every category — nothing filtered out.

Created by **Dahham Alsoud**

*Why not just use PubMed alerts? PubMed alerts give you a flat, unsorted list with no grouping by topic — you still have to manually scan through everything to find what's relevant to your work. IBD LitMonitor organizes the week's literature into clinical subfields so you can go straight to what matters, whether that's therapeutics, pediatrics, surgery, or any other domain.*
:::

---

### This Week's Coverage

::: {.grid}

::: {.g-col-3}
::: {.stat-box}
::: {.stat-number}
`r nrow(papers)`
:::
::: {.stat-label}
Total Papers
:::
:::
:::

::: {.g-col-3}
::: {.stat-box}
::: {.stat-number}
`r papers %>% distinct(category) %>% nrow()`
:::
::: {.stat-label}
Categories
:::
:::
:::

:::

---

## Papers by Category

::: {.callout-note icon=false}
## How to Read This Report

**Coverage:** `r nrow(get_search_queries() |> unlist())` PubMed queries across `r length(get_search_queries())` IBD domains, deduplicated — each paper appears exactly once in the most specific matching category.

**Categorisation:** Automatic and imperfect — a paper found by multiple queries is assigned to the most specific one (e.g. a vedolizumab trial in children → Pediatric IBD). Use the [dashboard](dashboard.html) to search across all categories.

**Study design badges** <span style="background:#2e7d32;color:#fff;font-size:0.72em;font-weight:600;padding:1px 6px;border-radius:3px;">RCT</span> <span style="background:#1565c0;color:#fff;font-size:0.72em;font-weight:600;padding:1px 6px;border-radius:3px;">Meta-analysis</span> <span style="background:#666;color:#fff;font-size:0.72em;font-weight:600;padding:1px 6px;border-radius:3px;">Review</span> — shown only when PubMed has explicitly assigned a publication type. Papers published within the last 1–2 weeks often don't have a badge yet.

**★ Flagship journals:** Papers from Gut, Gastroenterology, Lancet, NEJM, JCC, and other leading GI journals are marked ★.

**Spotted something wrong or have a suggestion?** &nbsp;<a href="mailto:dahhamalsoud@gmail.com?subject=IBD%20LitMonitor%20Feedback&body=Type%20of%20feedback%20(please%20keep%20one)%3A%0A%0A%5B%20%5D%20Miscategorised%20paper%0APMID%3A%0AShould%20be%20in%20category%3A%0A%0A%5B%20%5D%20Feature%20suggestion%0ADescription%3A%0A%0A%5B%20%5D%20Other%0ADescription%3A" style="display:inline-block;padding:3px 12px;background:#8b3a1e;color:#fff;border-radius:4px;text-decoration:none;font-size:0.85em;font-weight:600;">Send feedback →</a>
:::

*Click a category to expand. Click a paper title to read its abstract.*

```{r}
#| results: asis

category_counts <- papers %>% count(category, sort = TRUE)

for (cat_name in category_counts$category) {
  cat_papers <- papers %>%
    filter(category == cat_name) %>%
    arrange(desc(publication_date))

  # Outer fold: category
  cat(glue('\n<details>\n<summary><strong>{cat_name}</strong> &nbsp;<span style="color:#8b3a1e;font-weight:600;">({nrow(cat_papers)} papers)</span></summary>\n\n'))

  for (i in seq_len(nrow(cat_papers))) {
    paper <- cat_papers[i, ]

    abstract_full <- if (!is.na(paper$abstract) && nchar(str_squish(paper$abstract)) > 0) {
      str_squish(paper$abstract)
    } else "Abstract not available."

    doi_line <- if (!is.na(paper$doi) && paper$doi != "") {
      glue('<a href="https://doi.org/{paper$doi}" target="_blank">DOI: {paper$doi}</a> &nbsp;|&nbsp; ')
    } else ""

    # Study design badge — only shown when PubMed has explicitly assigned a type.
    # Nothing is shown when untagged (common for papers < 2 weeks old).
    design_badge <- if (!is.na(paper$study_design)) {
      badge_color <- switch(paper$study_design,
        "RCT"              = "#2e7d32",
        "Meta-analysis"    = "#1565c0",
        "Systematic review"= "#1565c0",
        "Clinical trial"   = "#388e3c",
        "Multicenter study"= "#558b2f",
        "#666"  # Review, Cohort, Case report, Editorial, Letter, etc.
      )
      glue('<span style="background:{badge_color};color:#fff;font-size:0.72em;font-weight:600;padding:1px 6px;border-radius:3px;margin-left:4px;">{paper$study_design}</span>')
    } else {
      ""   # no badge if PubMed hasn't tagged it yet
    }

    # Flagship journal marker
    flagship_badge <- if (isTRUE(paper$flagship)) {
      '<span style="color:#b8860b;font-size:0.8em;margin-left:4px;" title="High-impact IBD/GI journal">&#9733;</span>'
    } else ""

    # Inner fold: individual paper — title + journal/date in summary, full abstract in body
    cat(glue('
<details>
<summary>{paper$title}{design_badge}{flagship_badge} <span style="color:#666;font-size:0.85em;font-weight:normal;">— {paper$journal} &nbsp;|&nbsp; {paper$publication_date}</span></summary>

<div style="padding:0.6em 0.8em;background:#f9f9f9;border-left:3px solid #8b3a1e;margin:0.4em 0;">
<p style="margin:0 0 0.5em 0;color:#444;">{abstract_full}</p>
<p style="margin:0;font-size:0.85em;color:#666;"><em>{paper$authors}</em></p>
<p style="margin:0.5em 0 0 0;font-size:0.9em;">{doi_line}<a href="{paper$url}" target="_blank">View on PubMed →</a></p>
</div>

</details>
'))
  }

  cat('\n</details>\n\n---\n\n')
}
```

---

## Category Distribution

```{r}
#| fig-width: 10
#| fig-height: 6

papers %>%
  count(category, sort = TRUE) %>%
  ggplot(aes(x = reorder(category, n), y = n)) +
  geom_col(fill = "#8b3a1e") +
  geom_text(aes(label = n), hjust = -0.2, size = 4) +
  coord_flip() +
  labs(
    title = "Papers by Category This Week",
    x = NULL,
    y = "Number of Papers"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    plot.title = element_text(face = "bold", size = 16),
    panel.grid.major.y = element_blank()
  ) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.15)))
```

---

::: {.callout-tip}
## Want More Control?

- [Interactive Dashboard](dashboard.html) — Filter by category, drug, and keywords
- [Subscribe via RSS](rss.xml) — All categories in one feed
- [Per-category RSS feeds](rss/) — Subscribe to just one subfield (e.g. [Therapeutics](rss/therapeutics_mechanisms.xml), [Pediatric IBD](rss/pediatric_ibd.xml))
- [Download Data](data/papers.json) — JSON format for your own analysis
:::

